Exploring the conformational dynamics of alanine dipeptide in solution subjected to an external electric field: A nonequilibrium molecular dynamics simulation

### 2013-11-18 writedown/npt-neq/alanine.tex
my revise, according to Giovanni's notes

\newcommand{\recheck}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}
\newcommand{\bluec}[1]{{\color{blue} #1}}
\newcommand{\greenc}[1]{{\color{green} #1}}
\newcommand{\vect}[1]{\textbf{\textit{#1}}}
\newcommand{\dd}[1]{\textsf{#1}}
\newcommand{\fwd}[0]{\textrm{fwd}}
\newcommand{\bwd}[0]{\textrm{bwd}}

@@

\title{Exploring the conformational dynamics of alanine dipeptide in solution subjected to an external electric field: A nonequilibrium molecular dynamics simulation}
% \title{The non-equilibrium molecular dynamics simulation of the alanine dipeptide under electric field}
\author{Han Wang}
\email{han.wang@fu-berlin.de}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\author{Christof Sch\"utte}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\affiliation{Zuse Institute Berlin, Germany}
\author{Giovanni Ciccotti}
\affiliation{Department of Physics, University of Roma "La Sapienza", Italy}
\author{Luigi Delle Site}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}

@@

In the present paper, we study a small peptide, alanine dipeptide, as a model for showing how the method can be used. 
The choice of a small peptide, rather than a larger one, is done firstly because it
is easier to draw clear conclusions on the nonequilibrium properties
of a single peptide by excluding the slow-going interplay between
different peptide segments along a chain. 
Secondly, the computational cost required is rather massive and thus  testing
the various methodology proposed here would be hampered for large molecules,
although they remain our goal.  \recheck{In any case, alanine dipeptide
is large enough so that the conformational changes of its internal degrees of freedom represent valid observables for a test of validity and utility of our approach.}
The paper is organized as follows:
\recheck{
  We firstly describe the idea of D-NEMD, and propose a technical improvement consisting in the definition and application of a local thermostating process.
Next, after introducing the physical
observables chosen to analyze the nonequilibrium response,
the numerical results for the case of a static and oscillatory EF, respectively, are discussed.} By calculating the nonequilibrium properties from D-NEMD
simulations, we are able to address
a few relevant questions: (i)~How the conformation of the peptide
changes in an EF (conformational space);
(ii)~When these changes happen (time scale),
and how likely they are (probability). 
To our knowledge, this study is one of the few (if not the first)  where
time-dependent nonequilibrium w.r.t. conformation
is explicitly considered for a (relatively large) molecule
in solution subject to an external time-dependent EF.

\section{Methodology}
\subsection{Nonequilibrium molecular dynamics simulation}
In this section we briefly review the dynamical approach to nonequilibrium
molecular dynamics (D-NEMD) \cite{ciccotti1975direct, ciccotti1979thought,
  orlandini2011hydrodynamics, orlandini2011hydrodynamics-01}.
In the following we denote the macroscopic observable by $O(t)$. If at $t$
the configurational probability distribution is $\rho(\vect x, t)$, where
$\vect x$ is the phase space variable, then the observable can be
expressed by
\begin{align}\label{eqn:tmp1}
  O(t) = \int d\vect x\, \hat O(\vect x)\rho(\vect x, t)  = \langle \hat O(\vect x), \rho(\vect x, t)\rangle,
\end{align}
we refer to $\hat O (\vect x)$ as the microscopic observable, 
which is measured at each point in the phase space $\vect x$.
Here we always assume that the initial probability distribution
$\rho(\vect x, 0)$ is known. In particular, in our case,
it is identical to the equilibrium distribution of the system
without EF.
The bracket on the right hand side of Eq.~\eqref{eqn:tmp1} denotes, as usual, the inner product in the
phase space.  We assume that the dynamics of the system is governed by the
Hamiltonian equation, i.e. $\dot {\vect x} = J \cdot \nabla_{\vect x}
\mathcal H(\vect x, t)$, where $\mathcal H$ is the time-dependent Hamiltonian,
and $J$ is
the symplectic matrix. The Liouville equation for the probability
distribution writes:
\begin{align}\label{eqn:tmp2}
  \frac{\partial \rho(\vect x, t)}{\partial t} = - iL(t) \rho(\vect x, t),
\end{align}
where $iL(t) = \{\cdot, \mathcal H\}$ is the Liouville operator.
Eq.~\eqref{eqn:tmp2}
can be formally solved by $\rho(\vect x, t) = U^\dagger(t,0) \rho(\vect x, 0)$,
where $U^\dagger(t,0) = \mathcal T\exp\{-i \int_0^t d t' L(t')\}$, and $\mathcal T$ is the time ordering operator.
On the other hand,
\begin{align}
  \frac{d \hat O(\vect x(t))}{dt} = \nabla_{\vect x}\hat O\cdot \dot{\vect x}
  = \nabla_{\vect x}\hat O\cdot J\cdot \nabla_{\vect x}\mathcal H
  = iL(t) \hat O (\vect x(t)).
\end{align}
This equation can be formally solved by
$\hat O(\vect x(t)) = U(t,0)\,\hat O(\vect x(0))$, therefore,
\begin{align}\nonumber
  O(t) & = \langle \hat O(\vect x), \rho(\vect x, t)\rangle
  = \langle \hat O(\vect x), U^\dagger(t,0)\, \rho(\vect x, 0)\rangle
  = \langle U(t,0)\,\hat O(\vect x), \rho(\vect x, 0)\rangle\\\label{eqn:tmp4}
  &= \langle \hat O(\vect x(t)), \rho(\vect x, 0)\rangle.
\end{align}
Since we assume that the system starts from the equilibrium distribution (without EF)
 Eq.~\eqref{eqn:tmp4} expresses the fact that the 
observable $O(t)$ calculated in a situation of nonequilibrium is equal to the ensemble average of
the microscopic observable computed along trajectories, starting from
initial configurations sampled from an \emph{equilibrium}
distribution. In practical terms we proceed by first running an equilibrium MD simulation
to generate a sample of configurations. Next we employ these configurations as
initial configurations and, for each, the full Hamiltonian dynamics is integrated until
time $t$.
In this paper, these trajectories are called \emph{branching
  trajectories}.
% {\bf EITHER WE PUT THE PICTURE OR REMOVE THIS SENTENCE {(see Fig.\ref{branching} for a pictorial representation)}}. 
Finally, the value of the macroscopic observable at time $t$ is
calculated by averaging the microscopic observable measured at each time $t$ along each branching trajectory.
Stochastic dynamics (e.g.~Langevin dynamics) can be handled analogously.

@@

\subsection{System settings and simulation protocol}
The system is set up in a $2.7\times 2.7\times
2.7\, \textsf{nm}^3$ periodic simulation region, with one alanine dipeptide
described by the CHARMM27 force field, and dissolved in 644 TIP3P
water molecules.
%The EF exserts force $\vect F = q \vect E(t)$ on
%the partial charge $q$ in the system.
The size of the  dynamical region is $R_{ex} = 1.0$~nm.
All simulations are performed by a home-modified GROMACS 4.5~\cite{pronk2013gromacs}.
First, an equilibrium NVT simulation at 300~K of
100~\textsf{ns} was performed with a Langevin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$).  Along the trajectory, configurations were taken every
50~\textsf{ps} and we used 2000 initial configurations for each nonequilibrium
MD simulation (if not stated otherwise for specific cases).
The branching trajectories were integrated by the
Leap-frog scheme with the aforementioned nonequilibrium
temperature control technique.  The
time step was $\Delta t = 0.002~\textsf{ps}$. The short-range
interaction was smoothed from $0.8$ to $1.0~\textsf{nm}$ by the
``switch'' method provided by GROMACS.  An energy conserving PME
method was applied to calculate the electrostatic interaction in this
periodic system. In the thermostated region, the original dynamics was
coupled to a Langevin thermostat with $\tau_T = 0.1~\textsf{ps}$.
\recheck{
  In all testing cases, this local thermostat is able to control
  the system at the desired temperature, i.e.~300~K (the results
  are not presented in this paper).
}
The whole system is also coupled to a Parrinello-Rahman barostat~\cite{parrinello1981polymorphic} with $\tau_P = 2.0~\textsf{ps}$ to keep
the pressure at ambient condition (1~Bar). Since the
change of the system size is small and slow, the pressure control
does not have an sizable effect on the
dynamics of the system.
At time $t=0$~ps, the system
has been fully equilibrated without any EF. From $t=0$ to
the warm-up time $t=t_{\textrm{warm}}$, the EF is switched on linearly, while
after $t=t_{\textrm{warm}}$, the field is kept constant in time at
$E_{\infty}$.  In this work we consider $t_{\textrm{warm}} = 10$~ps,
and $E_{\infty} = 1$~V/nm.
The direction of the field is arbitrarily chosen along the
$x$ direction.
In this paper we denote the
EF as a function of time $\vect E(t)$.
Therefore, for the case of constant EF, we have
$\vect E(t) = (E_\infty\cdot t/t_{warm},0,0)$ for $0\leq t < t_{warm}$, and 
$\vect E(t) = (E_\infty,0,0)$
for $t \geq t_{warm}$.

@@

\begin{figure}
  \centering
  \includegraphics[width=1.00\textwidth]{fig-all.eps}
  % \raisebox{0.2\height}{\includegraphics[width=0.22\textwidth]{c-2.eps}}
  % \includegraphics[width=0.37\textwidth]{fig-begin-1.eps}
  % \includegraphics[width=0.37\textwidth]{fig-end-1.eps}
  \caption{
    \recheck{
    The probability density of the conformations of the alanine dipeptide plotted on the Ramachandran plot.
    (a): Two dihedral angles
    $\phi$ and $\psi$ are used to represent the molecular conformations of the alanine dipeptide.
    (b): The equilibrium probability density 
    at $t=0$~\textsf{ps}. (c): The probability density of equilibrium at $t=1000$~\textsf{ps}, resulting from the application of a constant EF of 1~V/nm. Darker color
    indicates higher probability.
    Plots are divided into 5 subregions corresponding
    to 5 molecular conformations:
    $A_1 = \{(\phi, \psi) | -134^\circ \leq \phi <  13^\circ, -125^\circ \leq \psi < 74^\circ\}$,
    $A_2 = \{(\phi, \psi) |  128^\circ \leq \phi <-134^\circ, -125^\circ \leq \psi < 74^\circ\}$,
    $B_1 = \{(\phi, \psi) | -110^\circ \leq \phi <  13^\circ,   74^\circ \leq \psi <-125^\circ\}$,
    $B_2 = \{(\phi, \psi) |  128^\circ \leq \phi <-110^\circ,   74^\circ \leq \psi <-125^\circ\}$,
    $C   = \{(\phi, \psi) |   13^\circ \leq \phi < 128^\circ, -180^\circ \leq \psi < 180^\circ\}$.
    Considering the angular periodicity, if we write, for example,
    $128^\circ \leq \phi <-134^\circ$, it implies that $\phi$ is in a range of $98$ degrees, between $128^\circ$ and $226^\circ$ ($226^\circ$ correspond to $-134^\circ$).
    This division is the same for both plot (b) and (c).
    }
  }
  \label{fig:tmp4}
\end{figure}

\recheck{
The change of conformation in time of the alanine dipeptide is
investigated by analyzing the probability density $p(\phi,\psi,t)$ on
the Ramachandran plot at time $t$. The definition of the dihedral
angles $\phi$ and $\psi$ is given in Fig.~\ref{fig:tmp4}~(a).
The probability density of conformations of equilibrium in absence of external field is shown in Fig.~\ref{fig:tmp4}~(b), while the same quantity, resulting from the molecular relaxation to the action of $E_{\infty} = 1$~V/nm is
given in Fig.~\ref{fig:tmp4}~(c). In the plot one can identify several clusters
in which conformations are grouped.
Therefore, in order to simplify the analysis, we divide the
Ramachandran plot into 5 subregions $\{A_1, A_2, B_1, B_2, C\}$
(see caption of Fig.~\ref{fig:tmp4} for the corresponding definition).
It is worth to notice that these conformations
are found in different secondary structures of a peptide chain:
$A_1$ and $A_2$ are
corresponding to the $\alpha$-helix. $B_1$ and $B_2$ are
corresponding to the $\beta$-sheet. $C$ is corresponding to the
left-handed $\alpha$-helix.
}

\recheck{From Fig.~\ref{fig:tmp4}, it is evident that the probability with which a
conformation occurs changes when the external EF is applied.
The most evident case is conformation $C$; in fact it is {almost} not present before the external field is applied but appears in a clear way as a response to the action of the EF. Following this line of thought, we are interested to study the change of conformation under the action of EF. Specifically, for each conformational change, we will analyze the relation between the structural relaxation of the molecule and its corresponding time scale. 
The starting point is the calculation of the probability
of each conformation}:
\begin{align}
  P_I(t) = \mathbb P (\vect x_t \in I), \quad  I \in \{A_1, A_2, B_1, B_2, C\},
\end{align}
Notice that the probability is time dependent, being an observable in nonequilibrium situation.
Next,
we consider the net probability flux from conformation $I$ to $J$, defined by:
\begin{align}\nonumber
  F_{J,I}(t) & = \frac1{\Delta t}\, [\mathbb P( \vect x_{t-\Delta t} \in J, \vect x_t \in I) - \mathbb P( \vect x_{t-\Delta t} \in I, \vect x_t \in J)], \\\label{eqn:tmp7}
  & J,I \in \{A_1, A_2, B_1, B_2, C\}.
\end{align}
\recheck{Where $\mathbb P( \vect x_{t-\Delta t} \in J, \vect x_t \in I)$
is the joint probability of the alanine being in conformation $J$ at
time $(t-\Delta t)$ \emph{and} being in conformation $I$ at time $t$.
The same definition applies to $\mathbb P( \vect x_{t-\Delta t} \in I, \vect x_t \in J)$.}
Here the time interval $\Delta t$ should be small enough compared to the
time scale of conformational dynamics,
so that the changes in the conformation probabilities
can be treated in linear approximation.
At the same time,  $\Delta t$ should be also large enough compared with
the timestep of the MD integrator, 
so that the quantities of Eq.~\eqref{eqn:tmp7} are well-defined, and can be
estimated with sufficient numerical accuracy (see also Ref.~\cite{schuette2011markov}).
A positive value \recheck{of $F_{J,I}(t)$} indicates that the
net flux goes from $J$ to $I$,
while a negative value indicates a net flux from $I$ to $J$.
Through $F_{J,I}(t)$, the analysis of the nonequilibrium process 
is projected onto the analysis of the probabilistic link between discretized conformations $\{A_1, A_2, B_1, B_2, C\}$.
\recheck{In the case of oscillatory EF, the behaviour of the net probability flux is also highly oscillatory, thus, in order to clearly identify the essential features of the molecular conformational changes, it is convenient to study the integrated net probability flux, defined as:
\begin{align}\label{eqn:tmp7a}
  Q_{J,I} (t) = \int_0^t F_{J,I}(\tau)d \tau,
\end{align}
which expresses the cumulative effects due to the action of EF on the conformations of the molecule.}

@@

\begin{figure}
  \centering
  \includegraphics[]{fig-meta-npt.eps}
  \caption{\recheck{The time-dependent probability of conformations $P_I(t)$ 
    when a constant EF is present, where $\sum_I P_I(t) = 1, \forall t$.}
    The warm-up time is $t_{warm} = 10$~ps. The red line refers to conformation $I = A_1$,
    green to $A_2$, dark blue to $B_1$, purple to $B_2$ and light blue
    to $C$. 
  }
  \label{fig:tmp5}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.3\textwidth]{fig-trans-010-flux-1.eps}
  \includegraphics[width=0.3\textwidth]{fig-trans-010-flux-2.eps}
  \includegraphics[width=0.3\textwidth]{fig-trans-010-flux-3.eps}
  \includegraphics[width=0.3\textwidth]{fig-trans-010-flux-4.eps}
  \includegraphics[width=0.3\textwidth]{fig-trans-010-flux-5.eps}
  \caption{\recheck{
    The net probability flux for the constant EF case.
    The warm-up time is $t_{warm} = 10$~ps.
    The probability flux $F_{J,I}(t)$ (defined by Eqn.~\eqref{eqn:tmp7}) is reported in units of $\textrm{ps}^{-1}$.
    From top, left to right we report $I = A_1$, $A_2$, $B_1$; bottom, left to right $B_2$ and
    $C$, respectively. In each plot, the red line stands for $J=A_1$,
    green for $J=A_2$, dark blue for $J=B_1$, purple for $J=B_2$ and light blue
    for $J=C$.} 
    }
  \label{fig:tmp6}
\end{figure}

>>>>

\newcommand{\recheck}[1]{{\color{red} #1}}
\newcommand{\redc}[1]{{\color{red} #1}}
\newcommand{\bluec}[1]{{\color{blue} #1}}
\newcommand{\greenc}[1]{{\color{green} #1}}
\newcommand{\vect}[1]{\textbf{\textit{#1}}}
\newcommand{\dd}[1]{\textsf{#1}}
\newcommand{\fwd}[0]{\textrm{fwd}}
\newcommand{\bwd}[0]{\textrm{bwd}}
\newcommand{\period}[0]{T_{\textrm{P}}}

@@

\title{Exploring the conformational dynamics of alanine dipeptide in solution subjected to an external electric field: A nonequilibrium molecular dynamics simulation}
% \title{The non-equilibrium molecular dynamics simulation of the alanine dipeptide under electric field}
\author{Han Wang}
\email{han.wang@fu-berlin.de}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\author{Christof Sch\"utte}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}
\affiliation{Zuse Institute Berlin, Germany}
\author{Giovanni Ciccotti}
% \affiliation{Department of Physics, University of Roma "La Sapienza", Italy}
\affiliation{School of Physics, Room 302B UCD-EMSC, University College Dublin, Belfield, Dublin 4, Ireland}
\affiliation{Dipartimento di Fisica, Università “La Sapienza” and CNISM, Piazzale Aldo Moro 5, 00185 Roma, Italy}
\author{Luigi Delle Site}
\affiliation{Institute for Mathematics, Freie Universit\"at Berlin, Germany}

@@

In the present paper, we study a small peptide, alanine dipeptide, as a model for showing how the method can be used. 
The choice of a small peptide, rather than a larger one, is done firstly because it
is easier to draw clear conclusions on the nonequilibrium properties
of a single peptide by excluding the slow-going interplay between
different peptide segments along a chain. 
Secondly, the computational cost required is rather massive and thus  testing
the various methodology proposed here would be hampered for large molecules,
although they remain our goal.  In any case, alanine dipeptide
is large enough so that the conformational changes of its internal degrees of freedom represent valid observables for a test of validity and utility of our approach.
The paper is organized as follows:
  We firstly describe the idea of D-NEMD, and propose a technical improvement consisting in the definition and application of a local thermostating process.
Next, after introducing the physical
observables chosen to analyze the nonequilibrium response,
the numerical results for the case of a static and oscillatory EF, respectively, are discussed. By calculating the nonequilibrium properties from D-NEMD
simulations, we are able to address
a few relevant questions: (i)~How the conformation of the peptide
changes in an EF (conformational space);
(ii)~When these changes happen (time scale),
and how likely they are (probability). 
To our knowledge, this study is one of the few (if not the first)  where
time-dependent nonequilibrium w.r.t. conformation
is explicitly considered for a (relatively large) molecule
in solution subject to an external time-dependent EF.

\section{Methodology}
\subsection{Nonequilibrium molecular dynamics simulation}
In this section we briefly review the dynamical approach to nonequilibrium
molecular dynamics (D-NEMD) \cite{ciccotti1975direct, ciccotti1979thought,
  orlandini2011hydrodynamics, orlandini2011hydrodynamics-01}.
In the following we denote the macroscopic observable by $O(t)$. If at $t$
the configurational probability distribution is $\rho(\vect x, t)$, where
$\vect x$ is the phase space variable, then the observable can be
expressed by
\begin{align}\label{eqn:tmp1}
  O(t) = \int d\vect x\, \hat O(\vect x)\rho(\vect x, t)  = \langle \hat O(\vect x), \rho(\vect x, t)\rangle,
\end{align}
we refer to $\hat O (\vect x)$ as the microscopic observable, 
which is measured at each point in the phase space $\vect x$.
Here we always assume that the initial probability distribution
$\rho(\vect x, 0)$ is known. In particular, in our case,
it is identical to the equilibrium distribution of the system
without EF.
The bracket on the right hand side of Eq.~\eqref{eqn:tmp1} denotes, as usual, the inner product in the
phase space.  We assume that the dynamics of the system is governed by the
Hamiltonian equation, i.e. $\dot {\vect x} = J \cdot \nabla_{\vect x}
\mathcal H(\vect x, t)$, where $\mathcal H$ is the time-dependent Hamiltonian,
and $J$ is
the symplectic matrix. The Liouville equation for the probability
distribution writes:
\begin{align}\label{eqn:tmp2}
  \frac{\partial \rho(\vect x, t)}{\partial t} = - iL(t) \rho(\vect x, t),
\end{align}
where $iL(t) = \{\cdot, \mathcal H\}$ is the Liouville operator.
Eq.~\eqref{eqn:tmp2}
can be formally solved by $\rho(\vect x, t) = U^\dagger(t,0) \rho(\vect x, 0)$,
\recheck{where $U^\dagger(t,0) = \mathcal T\exp\{-i \int_0^t d t' L(t')\}$, and $\mathcal T$ is the time ordering operator.}
On the other hand,
\begin{align}
  \frac{d \hat O(\vect x(t))}{dt} = \nabla_{\vect x}\hat O\cdot \dot{\vect x}
  = \nabla_{\vect x}\hat O\cdot J\cdot \nabla_{\vect x}\mathcal H
  = iL(t) \hat O (\vect x(t)).
\end{align}
This equation can be formally solved by
$\hat O(\vect x(t)) = U(t,0)\,\hat O(\vect x(0))$, therefore,
\begin{align}\nonumber
  O(t) & = \langle \hat O(\vect x), \rho(\vect x, t)\rangle
  = \langle \hat O(\vect x), U^\dagger(t,0)\, \rho(\vect x, 0)\rangle
  = \langle U(t,0)\,\hat O(\vect x), \rho(\vect x, 0)\rangle\\\label{eqn:tmp4}
  &= \langle \hat O(\vect x(t)), \rho(\vect x, 0)\rangle.
\end{align}
Since we assume that the system starts from the equilibrium distribution (without EF)
 Eq.~\eqref{eqn:tmp4} expresses the fact that the 
observable $O(t)$ calculated in a situation of nonequilibrium is equal to the ensemble average of
the microscopic observable computed along trajectories, starting from
initial configurations sampled from an \emph{equilibrium}
distribution. In practical terms we proceed by first running an equilibrium MD simulation
to generate a sample of configurations. Next we employ these configurations as
initial configurations and, for each, the full Hamiltonian dynamics is integrated until
time $t$.
In this paper, these trajectories are called \emph{branching
  trajectories}.
% {\bf EITHER WE PUT THE PICTURE OR REMOVE THIS SENTENCE {(see Fig.\ref{branching} for a pictorial representation)}}. 
Finally, the value of the macroscopic observable at time $t$ is
calculated by averaging the microscopic observable measured at each time $t$ along each branching trajectory.
Stochastic dynamics (e.g.~Langevin dynamics) can be handled analogously.

@@

\subsection{System settings and simulation protocol}
The system is set up in a $2.7\times 2.7\times
2.7\, \textsf{nm}^3$ periodic simulation region, with one alanine dipeptide
described by the CHARMM27 force field, and dissolved in 644 TIP3P
water molecules.
%The EF exserts force $\vect F = q \vect E(t)$ on
%the partial charge $q$ in the system.
The size of the  dynamical region is $R_{ex} = 1.0$~nm.
All simulations are performed by a home-modified GROMACS 4.5~\cite{pronk2013gromacs}.
First, an equilibrium NVT simulation at 300~K of
100~\textsf{ns} was performed with a Langevin thermostat (time-scale
$\tau_T = 0.5~\textsf{ps}$).  Along the trajectory, configurations were taken every
50~\textsf{ps} and we used 2000 initial configurations for each nonequilibrium
MD simulation (if not stated otherwise for specific cases).
The branching trajectories were integrated by the
Leap-frog scheme with the aforementioned nonequilibrium
temperature control technique.  The
time step was $\Delta t = 0.002~\textsf{ps}$. The short-range
interaction was smoothed from $0.8$ to $1.0~\textsf{nm}$ by the
``switch'' method provided by GROMACS.  An energy conserving PME
method was applied to calculate the electrostatic interaction in this
periodic system. In the thermostated region, the original dynamics was
coupled to a Langevin thermostat with $\tau_T = 0.1~\textsf{ps}$.
  In all testing cases, this local thermostat is able to control
  the system at the desired temperature, i.e.~300~K (the results
  are not presented in this paper).
The whole system is also coupled to a Parrinello-Rahman barostat~\cite{parrinello1981polymorphic} with $\tau_P = 2.0~\textsf{ps}$ to keep
the pressure at ambient condition (1~Bar). Since the
change of the system size is small and slow, the pressure control
does not have an sizable effect on the
dynamics of the system.
At time $t=0$~ps, the system
has been fully equilibrated without any EF. From $t=0$ to
the warm-up time $t=t_{\textrm{warm}}$, the EF is switched on linearly, while
after $t=t_{\textrm{warm}}$, the field is kept constant in time at
$E_{\infty}$.  In this work we consider $t_{\textrm{warm}} = 10$~ps,
and $E_{\infty} = 1$~V/nm.
The direction of the field is arbitrarily chosen along the
$x$ direction.
In this paper we denote the
EF as a function of time $\vect E(t)$.
Therefore, for the case of constant EF, we have
$\vect E(t) = (E_\infty\cdot t/t_{warm},0,0)$ for $0\leq t < t_{warm}$, and 
$\vect E(t) = (E_\infty,0,0)$
for $t \geq t_{warm}$.

@@

\begin{figure}
  \centering
  \includegraphics[width=1.00\textwidth]{fig-all.eps}
  % \raisebox{0.2\height}{\includegraphics[width=0.22\textwidth]{c-2.eps}}
  % \includegraphics[width=0.37\textwidth]{fig-begin-1.eps}
  % \includegraphics[width=0.37\textwidth]{fig-end-1.eps}
  \caption{
    The probability density of the conformations of the alanine dipeptide plotted on the Ramachandran plot.
    (a): Two dihedral angles
    $\phi$ and $\psi$ are used to represent the molecular conformations of the alanine dipeptide.
    (b): The equilibrium probability density 
    at $t=0$~\textsf{ps}. (c): The probability density of equilibrium at $t=1000$~\textsf{ps}, resulting from the application of a constant EF of 1~V/nm. Darker color
    indicates higher probability.
    Plots are divided into 5 subregions corresponding
    to 5 molecular conformations:
    $A_1 = \{(\phi, \psi) | -134^\circ \leq \phi <  13^\circ, -125^\circ \leq \psi < 74^\circ\}$,
    $A_2 = \{(\phi, \psi) |  128^\circ \leq \phi < 226^\circ, -125^\circ \leq \psi < 74^\circ\}$,
    $B_1 = \{(\phi, \psi) | -110^\circ \leq \phi <  13^\circ,   74^\circ \leq \psi < 235^\circ\}$,
    $B_2 = \{(\phi, \psi) |  128^\circ \leq \phi < 250^\circ,   74^\circ \leq \psi < 235^\circ\}$,
    $C   = \{(\phi, \psi) |   13^\circ \leq \phi < 128^\circ, -180^\circ \leq \psi < 180^\circ\}$.
    \recheck{Considering the angular periodicity, The angle, for example, $226^\circ$ corresponds to $-134^\circ$
    on the plot (b) and (c).}
    % Considering the angular periodicity, if we write, for example,
    % $128^\circ \leq \phi <-134^\circ$, it implies that $\phi$ is in a range of $98$ degrees, between $128^\circ$ and $226^\circ$ ($226^\circ$ correspond to $-134^\circ$).
    This division is the same for both plot (b) and (c).
  }
  \label{fig:tmp4}
\end{figure}

The change of conformation in time of the alanine dipeptide is
investigated by analyzing the probability density $p(\phi,\psi,t)$ on
the Ramachandran plot at time $t$. The definition of the dihedral
angles $\phi$ and $\psi$ is given in Fig.~\ref{fig:tmp4}~(a).
The probability density of conformations of equilibrium in absence of external field is shown in Fig.~\ref{fig:tmp4}~(b), while the same quantity, resulting from the molecular relaxation to the action of $E_{\infty} = 1$~V/nm is
given in Fig.~\ref{fig:tmp4}~(c). In the plot one can identify several clusters
in which conformations are grouped.
Therefore, in order to simplify the analysis, we divide the
Ramachandran plot into 5 subregions $\{A_1, A_2, B_1, B_2, C\}$
(see caption of Fig.~\ref{fig:tmp4} for the corresponding definition).
It is worth to notice that these conformations
are found in different secondary structures of a peptide chain:
$A_1$ and $A_2$ are
corresponding to the $\alpha$-helix. $B_1$ and $B_2$ are
corresponding to the $\beta$-sheet. $C$ is corresponding to the
left-handed $\alpha$-helix.

From Fig.~\ref{fig:tmp4}, it is evident that the probability with which a
conformation occurs changes when the external EF is applied.
The most evident case is conformation $C$; in fact it is {almost} not present before the external field is applied but appears in a clear way as a response to the action of the EF. Following this line of thought, we are interested to study the change of conformation under the action of EF. Specifically, for each conformational change, we will analyze the relation between the structural relaxation of the molecule and its corresponding time scale. 
The starting point is the calculation of the probability
of each conformation:
\recheck{
\begin{align}
  P_I(t) = \mathbb P ( (\phi_t,\psi_t) \in I), \quad  I \in \{A_1, A_2, B_1, B_2, C\},
\end{align}
}
Notice that the probability is time dependent, being an observable in nonequilibrium situation.
Next,
we consider the net probability flux from conformation $I$ to $J$, defined by:
\recheck{
\begin{align}\nonumber
  F_{J,I}(t) & = \frac1{\Delta t}\,
  \bigg[\,
  \mathbb P \big( (\phi_{t-\Delta t},\psi_{t-\Delta t}) \in J,\, (\phi_t, \psi_t) \in I \,\big) 
  -
  \mathbb P \big( (\phi_{t-\Delta t}, \psi_{t-\Delta t}) \in I,\, (\phi_t, \psi_t) \in J \,\big)
  \bigg], \\\label{eqn:tmp7}
  & J,I \in \{A_1, A_2, B_1, B_2, C\}.
\end{align}}
\recheck{\textit{I cannot read what Giovanni wrote on the top of page 9.}}
Where $\mathbb P\big( (\phi_{t-\Delta t}, \psi_{t-\Delta t}) \in J,\, (\phi_t, \psi_t) \in I\,\big)$
is the joint probability of the alanine being in conformation $J$ at
time $(t-\Delta t)$ \emph{and} being in conformation $I$ at time $t$.
The same definition applies to
$\mathbb P \big( (\phi_{t-\Delta t}, \psi_{t-\Delta t}) \in I,\, (\phi_t, \psi_t) \in J\,\big)$.
Here the time interval $\Delta t$ should be small enough compared to the
time scale of conformational dynamics,
so that the changes in the conformation probabilities
can be treated in linear approximation.
At the same time,  $\Delta t$ should be also large enough compared with
the timestep of the MD integrator, 
so that the quantities of Eq.~\eqref{eqn:tmp7} are well-defined, and can be
estimated with sufficient numerical accuracy (see also Ref.~\cite{schuette2011markov}).
A positive value of $F_{J,I}(t)$ indicates that the
net flux goes from $J$ to $I$,
while a negative value indicates a net flux from $I$ to $J$.
Through $F_{J,I}(t)$, the analysis of the nonequilibrium process 
is projected onto the analysis of the probabilistic link between discretized conformations $\{A_1, A_2, B_1, B_2, C\}$.
In the case of oscillatory EF, the behaviour of the net probability flux is also highly oscillatory, thus, in order to clearly identify the essential features of the molecular conformational changes, it is convenient to study the integrated net probability flux, defined as:
\begin{align}\label{eqn:tmp7a}
  Q_{J,I} (t) = \int_0^t F_{J,I}(\tau)d \tau,
\end{align}
which expresses the cumulative effects due to the action of EF on the conformations of the molecule.

@@

\begin{figure}
  \centering
  \includegraphics[]{fig-meta-npt.eps}
  \caption{The time-dependent probability of conformations $P_I(t)$ 
    when a constant EF is present, where $\sum_I P_I(t) = 1, \forall t$.
    The warm-up time is $t_{warm} = 10$~ps. The red line refers to conformation $I = A_1$,
    green to $A_2$, dark blue to $B_1$, purple to $B_2$ and light blue
    to $C$. 
  }
  \label{fig:tmp5}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.3\textwidth]{fig-trans-010-flux-1.eps}
  \includegraphics[width=0.3\textwidth]{fig-trans-010-flux-2.eps}
  \includegraphics[width=0.3\textwidth]{fig-trans-010-flux-3.eps}
  \includegraphics[width=0.3\textwidth]{fig-trans-010-flux-4.eps}
  \includegraphics[width=0.3\textwidth]{fig-trans-010-flux-5.eps}
  \caption{
    The net probability flux for the constant EF case.
    The warm-up time is $t_{warm} = 10$~ps.
    The probability flux $F_{J,I}(t)$ (defined by Eqn.~\eqref{eqn:tmp7}) is reported in units of $\textrm{ps}^{-1}$.
    From top, left to right we report $I = A_1$, $A_2$, $B_1$; bottom, left to right $B_2$ and
    $C$, respectively. In each plot, the red line stands for $J=A_1$,
    green for $J=A_2$, dark blue for $J=B_1$, purple for $J=B_2$ and light blue
    for $J=C$.
    }
  \label{fig:tmp6}
\end{figure}

The time-dependent probability density of the system to stay in a certain
conformation is given in
Fig.~\ref{fig:tmp5}. The $\beta$-sheet conformations $B_1$ and $B_2$,
which are present at the initial stage after EF is switched on,
fade away as the system relaxes to the new equilibrium. The
probability of the $\alpha$-helix conformation $A_1$, grows from 0.45 to 0.69.
The $\alpha$-helix $A_2$  does not change significantly  under the action of EF.
The left-hand helix conformation $C$ noticeably grows from 0 to 0.22.
We plot the net probability fluxes in Fig.~\ref{fig:tmp6}.
Results show that for short $t$, the net fluxes are generally non-zero, however, 
as $t$ goes to infinity,
all fluxes converge to zero. This indicates
that as the EF is switched on, the system first is driven away from the initial
equilibrium (at zero EF);  This is a nonequilibrium process.
After a time long enough, the system
is fully relaxed to the new equilibrium defined by the action of the constant EF,
and stays there as long as the EF is kept.
An important point is that despite the system is finally relaxed to the new conformation,
the relaxation timescales corresponding to each conformation are rather different. 
From Fig.~\ref{fig:tmp5} and \ref{fig:tmp6}, we observe mainly
three different timescales 10~ps, 100~ps and 500~ps,
whose corresponding conformations are summarized in Tab.~\ref{tab:tmp1}.
This is a relevant result because it shows the possibility of employing an external EF as a tool to identify relevant time scales in the conformational behaviour of a molecule in solution.
